---
title: 'Assignment #4'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
library(dbplyr)            # for SQL query "cheating" - part of tidyverse but needs to be loaded separately
library(mdsr)              # for accessing some databases - goes with Modern Data Science with R textbook
library(RMySQL)            # for accessing MySQL databases
library(RSQLite)           # for accessing SQLite databases
theme_set(theme_minimal()) # Lisa's favorite theme
```

When you finish the assignment, remove the `#` from the options chunk at the top, so that messages and warnings aren't printed. If you are getting errors in your code, add `error = TRUE` so that the file knits. I would recommend not removing the `#` until you are completely finished.

## Put it on GitHub!        

[gitHub](https://github.com/yzhong0620/STAT-456-Assignment-4)

## SQL

### **Task 1**:

```{r}
con_air <- dbConnect_scidb("airlines")
```

```{r}
dataset <- tbl(con_air, "flights") %>% 
  group_by(origin, year, month) %>% 
  summarize(total_depart = n(),
            distinct_dest = count(distinct(dest)),
            ave_length = mean(arr_delay),
            ave_distance = mean(distance),
            prop_late = mean(arr_delay > 20)) %>% 
  ungroup() %>% 
  inner_join(tbl(con_air, "airports"),
             by = c("origin" = "faa")) %>% 
  select(name, year, month, total_depart, distinct_dest, ave_length, ave_distance, prop_late)
```

```{r}
show_query(dataset)
```

```{sql connection=con_air}
SELECT `name`, `year`, `month`, `total_depart`, `distinct_dest`, `ave_length`, `ave_distance`, `prop_late`
FROM (SELECT `LHS`.`carrier` AS `carrier`, `year`, `month`, `total_depart`, `distinct_dest`, `ave_length`, `ave_distance`, `prop_late`, `name`
FROM (SELECT `carrier`, `year`, `month`, COUNT(*) AS `total_depart`, count(distinct(`dest`)) AS `distinct_dest`, AVG(`arr_delay`) AS `ave_length`, AVG(`distance`) AS `ave_distance`, AVG(`arr_delay` > 20.0) AS `prop_late`
FROM `flights`
GROUP BY `carrier`, `year`, `month`) `LHS`
INNER JOIN `carriers` AS `RHS`
ON (`LHS`.`carrier` = `RHS`.`carrier`)
) `q01`
```

```{r, width=15}
dataset %>% 
  group_by(name) %>%
  summarise(ave_prop = mean(prop_late),
            ave_distance = round(mean(ave_distance), digits = 2)) %>%
  arrange(desc(ave_prop)) %>% 
  head(10) %>% 
  ggplot(aes(x = ave_prop, y = fct_reorder(name, ave_prop))) +
  geom_col(aes(fill = ave_distance)) +
  geom_label(aes(label = ave_distance)) +
  labs(x = "Proportion of flights that arrived more than 20 minutes late", y = "", fill = "Average distance of flights")
```

```{r}
library(gt)

dataset %>% 
  group_by(name) %>% 
  summarise(ave_prop = mean(prop_late),
            ave_distance = mean(ave_distance)) %>%
  arrange(desc(ave_prop)) %>% 
  head(4) %>% 
  gt() %>% 
  tab_header(
    title = md("**The worst airports in terms of late arrivals**")) %>% 
  fmt_number(
    columns = vars(ave_prop),
    decimals = 2) %>% 
  fmt_number(
    columns = vars(ave_distance),
    decimals = 2) %>%
  data_color(columns = vars(ave_distance),
             colors =  scales::col_numeric(palette = "viridis", 
                                           domain = NULL),
             alpha = .7) %>% 
  tab_options(column_labels.background.color = "gray")
```

### **Task 2**: 

```{r}
popularity <- tbl(con_air, "flights") %>% 
  group_by(origin, dest) %>% 
  summarise(route = paste(origin, dest, sep="->"),
            total = n(),
            prop_cancel = round(mean(cancelled), digits = 4)) %>% 
  ungroup() %>% 
  select(route, total, prop_cancel)
```

```{r}
show_query(popularity)
```

```{sql connection=con_air}
SELECT `route`, `total`, `prop_cancel`
FROM (SELECT `origin`, `dest`, CONCAT_WS('->', `origin`, `dest`) AS `route`, COUNT(*) AS `total`, AVG(`cancelled`) AS `prop_cancel`
FROM `flights`
GROUP BY `origin`, `dest`) `q01`
```

```{r}
popularity %>% 
  arrange(desc(total)) %>% 
  head(10) %>% 
  ggplot(aes(x = total, y = fct_reorder(route, total))) +
  geom_col(aes(fill = prop_cancel)) +
  geom_label(aes(label = prop_cancel)) +
  labs(x = "Total count of the route", y = "", fill = "Proportion of flights cancelled")
```

> What are the most popular routes and the proportion of flights cancelled in these routes?

## Shiny app

[gitHub](https://github.com/yzhong0620/shiny)

[app](https://yunyang-zhong.shinyapps.io/shiny/)
